Lies, versus the lying liars who tell them: comment on Garbarino, Slonim and Villeval (2018)

### 2018-11-01 GSV-comment-brief.Rmd
Update comment to optionally use proportion of lies as ground truth

---
title: "Lies, versus the lying liars who tell them: comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
output:
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  bookdown::tufte_book2:
    citation_package: natbib
    latex_engine: xelatex
    md_extensions: +raw_attribute
bibliography: bibliography.bib
link-citations: yes

---

@@

nreps <- 1000L
rewrite_csv <- FALSE # set to TRUE to (re)create the simulations CSV file
cache_results <- TRUE # FALSE to recalculate all results. TRUE for speed.
```

@@

simulate_cointoss <- function (row) {
  row["Liars"] + sum(rbinom(row["N"] - row["Liars"], 1, 1 - row["P"]))
}

@@

sim_data <- true_data %>%  
      cbind(as_tibble(results)) %>% 
      tidyr::gather("sim", "R", starts_with("V")) %>% 
      mutate(
        sim = as.numeric(sub("^V", "", sim)),
      ) %>% 
      arrange(N, Liars, P)

@@

# reshape to long by method
sim_est %<>%
      tidyr::gather(measure, value, contains(".")) %>% 
      tidyr::extract(measure, into = c("statistic", "method"), regex = "(.*)\\.(.*)") %>% 
      tidyr::spread(statistic, value)

sim_est %<>% mutate(
        true_prop = Liars/N,
        within_ci = true_prop >= LB & true_prop <= UB,
        error     = EV - true_prop,
        error_sq  = error^2
      )

@@

coverage <- sim_est %>% 
      group_by(method, CI) %>% 
      summarize(mean_in_ci = mean(within_ci))

@@

```{r disagg-coverage}
disagg <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = R/N, N, P, CI
      ) %>% 
      summarize(mean_in_ci = mean(within_ci), prop = n()/nrow(.)) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci) %>% 
      mutate(ok  = `CI 95%` > 0.95)

@@

```{r coverage-rn}
coverage_rn <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = factor(Hmisc::cut2(R/N, cuts = c(0, 1/2, 1)), labels = c("< 0.5", "&ge; 0.5"))      
      ) %>% 
      mutate(pct_cases = n()/nrow(.)) %>% 
      group_by(`R/N:`, CI) %>% 
      summarize(mean_in_ci = mean(within_ci), `Pct cases` = pct_cases[1]) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci)

@@

coverage_N <- sim_est %>% 
      filter(N <= 50) %>% 
      group_by(method, CI, N) %>% 
      summarize(mean_in_ci = mean(within_ci)) %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI ", CI, "%"),
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      select(Method, N, CI, mean_in_ci) %>% 
      spread(CI, mean_in_ci) %>% 
      arrange(factor(Method, levels = c("GSV", "Frequentist", "Bayesian")))

@@

sim_est %>% 
      filter(CI == 90) %>% 
      group_by(method, N) %>% 
      summarize(mean_sq_error = mean(error_sq)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("(#tab:point-estimates) Mean squared errors by method and N") %>% 
      theme_article()
```

@@

@@

coverage_recalc <- sim_est_recalc %>% 
      filter(method == "recalc") %>% 
      group_by(CI) %>% 
      summarize(mean_in_ci = mean(within_ci)) %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI: ", CI, "%"),
      ) %>% 
      spread(CI, mean_in_ci)

@@

coverage_treat_gsv <- sim_est %>% 
      filter(method == "gsv") %>% 
      group_by(CI, N, P) %>% 
      summarize(mean_in_ci = mean(within_ci)) %>% 
      ungroup() %>% 
      mutate(
        P = factor(paste0(P, "%")),
        ci = CI,
        CI = paste0(CI, "%")
      )

@@

coverage_grp_gsv <- sim_est %>% 
      filter(method == "gsv") %>% 
      group_by(CI, N, Liars, P) %>% 
      summarize(mean_in_ci = mean(within_ci)) %>% 
      ungroup() %>% 
      mutate(
        P = factor(paste0(P, "%")),
        ci = CI,
        CI = paste0(CI, "%")
      )

@@

```{r errors-by-true-prop, fig.cap = "Errors by method and true proportion of liars, N &le; 50"}
errors_true_prop <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(method, true_prop) %>% 
      summarize(mean_sq_error = mean(error_sq)) %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`True L/N` = true_prop, Method = method)

ggplot(errors_true_prop, aes(`True L/N`, mean_sq_error, colour = Method)) + 
      geom_line() + 
      geom_point() +
      ylab("Mean square error") + 
      scale_x_continuous(breaks = 0:5/5)
```

@@

errors_disagg <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(rn = findInterval(R/N, 0:10/10, rightmost.closed = TRUE), N, P, method) %>% 
      summarize(mean_sq_error = mean(error_sq)) %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`R/N` = rn, Method = method)

ggplot(errors_disagg, aes(`R/N`, mean_sq_error, colour = Method)) + 
      geom_line() + 
      geom_point() +
      ylab("Mean square error") + facet_wrap(~N + P, scales = "free_y")
```

>>>>

---
title: "Lies, versus the lying liars who tell them: comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
output:
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  bookdown::tufte_book2:
    citation_package: natbib
    latex_engine: xelatex
    md_extensions: +raw_attribute
bibliography: bibliography.bib
link-citations: yes
editor_options: 
  chunk_output_type: console
---

@@

nreps <- 1000L
rewrite_csv <- FALSE # set to TRUE to (re)create the simulations CSV file
cache_results <- TRUE # FALSE to recalculate all results. TRUE for speed.
estimate_lies_not_liars <- FALSE # if FALSE, use Liars/N as ground truth
                                # if TRUE, use (R-heads)/(N-heads)
```

@@

simulate_cointoss <- function (row) {
  h_liars <- rbinom(1, row["Liars"], 1 - row["P"])
  h_honest <- rbinom(1, row["N"] - row["Liars"], 1 - row["P"])
  R <- row["Liars"] + h_honest
  heads <- h_honest + h_liars
  c(R, heads)
}

@@

results_R <- results[1, , ]
results_heads <- results[2, , ]
colnames(results_R) <- paste0("R", 1:nreps)
colnames(results_heads) <- paste0("heads", 1:nreps)

sim_data <- true_data %>% 
      cbind(as_tibble(results_R)) %>% 
      cbind(as_tibble(results_heads)) %>% 
      gather("stat", "result", starts_with("R"), starts_with("heads")) %>%
      extract(stat, c("stat", "sim"), regex="(.*?)(\\d+)", convert = TRUE) %>%
      spread(stat, result) %>%  
      arrange(N, Liars, P)

@@

# reshape to long by method
sim_est %<>%
      tidyr::gather(measure, value, contains(".")) %>% 
      tidyr::extract(measure, into = c("statistic", "method"), regex = "(LB|EV|UB)\\.(.*)") %>% 
      tidyr::spread(statistic, value)

# alternative "ground truth" is proportion of lies: (R - heads)/(N - heads)
if (estimate_lies_not_liars) {
  sim_est %<>% mutate(
          true_prop = (R - heads)/(N - heads)
        )
} else {
  sim_est %<>% mutate(
          true_prop = Liars/N
        )
}

sim_est %<>% mutate(
        within_ci = true_prop >= LB & true_prop <= UB,
        error     = EV - true_prop,
        error_sq  = error^2
      )

@@

coverage <- sim_est %>% 
      group_by(method, CI) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE))

@@

```{r disagg-coverage}
disagg <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = R/N, N, P, CI
      ) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE), prop = n()/nrow(.)) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci) %>% 
      mutate(ok  = `CI 95%` > 0.95)

@@

```{r coverage-rn}
coverage_rn <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = factor(Hmisc::cut2(R/N, cuts = c(0, 1/2, 1)), labels = c("< 0.5", "&ge; 0.5"))      
      ) %>% 
      mutate(pct_cases = n()/nrow(.)) %>% 
      group_by(`R/N:`, CI) %>% 
      summarize(
        mean_in_ci  = mean(within_ci, na.rm = TRUE), 
        `Pct cases` = pct_cases[1]
      ) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci)

@@

coverage_N <- sim_est %>% 
      filter(N <= 50) %>% 
      group_by(method, CI, N) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI ", CI, "%"),
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      select(Method, N, CI, mean_in_ci) %>% 
      spread(CI, mean_in_ci) %>% 
      arrange(factor(Method, levels = c("GSV", "Frequentist", "Bayesian")))

@@

sim_est %>% 
      filter(CI == 90) %>% 
      group_by(method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("(#tab:point-estimates) Mean squared errors by method and N") %>% 
      theme_article()
```

@@

This approach has a nice interpretation, somewhat analogous
to the use of the null hypothesis in frequentist methods. We start from the 
belief that the proportion of liars is the same across all groups. The posterior
then shows how much the evidence changes this belief.

@@

coverage_recalc <- sim_est_recalc %>% 
      filter(method == "recalc") %>% 
      group_by(CI) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI: ", CI, "%"),
      ) %>% 
      spread(CI, mean_in_ci)

@@

coverage_treat_gsv <- sim_est %>% 
      filter(method == "gsv") %>% 
      group_by(CI, N, P) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        P = factor(paste0(P, "%")),
        ci = CI,
        CI = paste0(CI, "%")
      )

@@

coverage_grp_gsv <- sim_est %>% 
      filter(method == "gsv") %>% 
      group_by(CI, N, Liars, P) %>% 
      summarize(mean_in_ci = mean(within_ci, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        P = factor(paste0(P, "%")),
        ci = CI,
        CI = paste0(CI, "%")
      )

@@

```{r errors-by-true-prop, fig.cap = "Errors by method and true proportion of liars, N &le; 50"}
errors_true_prop <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(method, true_prop) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`True L/N` = true_prop, Method = method)

g <- ggplot(errors_true_prop, aes(`True L/N`, mean_sq_error, colour = Method)) + 
      ylab("Mean square error") + 
      scale_x_continuous(breaks = 0:5/5)

# proportion of liars is set to 0:10/10, but of lies is random,
# so we have to use a smoothing line
if (estimate_lies_not_liars) g + geom_smooth(method="loess") else 
      g + geom_line() + geom_point()

@@

errors_disagg <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(rn = findInterval(R/N, 0:10/10, rightmost.closed = TRUE), N, P, method) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`R/N` = rn, Method = method)

g <- ggplot(errors_disagg, aes(`R/N`, mean_sq_error, colour = Method)) + 
      ylab("Mean square error") + facet_wrap(~N + P, scales = "free_y")

if (estimate_lies_not_liars) g + geom_smooth(method="loess") else 
      g + geom_line() + geom_point()
```
